\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2021-11-17 doc/USER_MANUAL/03_mesh_generation.tex
renames input parameter subamp_seismos to NTSTEP_BETWEEN_OUTPUT_SAMPLE (making naming more uniform across different SPECFEM codes)

When you are ready to compile, in the main directory type
{\small
\begin{verbatim}
make xdecompose_mesh
\end{verbatim}
}
\noindent
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
should be produced.

@@

In case you successfully ran the configuration script, you are also
ready to compile the internal mesher. This is an alternative to CUBIT
for the mesh generation of relatively simple models.
%
\begin{figure}[htbp]
\begin{centering}
\includegraphics[scale=0.5]{figures/socal_map_mpi.pdf}
\par
\end{centering}
\caption{For parallel computing purposes,
the model block is subdivided in $\nprocxi\times\nproceta$ slices
of elements. In this example we use $5^{2}=25$ processors. }
\label{fig:For-parallel-computing}
\end{figure}

\noindent
In the main directory, type
{\small
\begin{verbatim}
make xmeshfem3D
\end{verbatim}
}
\noindent
If all paths and flags have been set correctly, the mesher should
now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
note that \texttt{xmeshfem3D} must be called directly from the main
directory, as most of the binaries of the package.

Input for the mesh generation program is provided through the parameter
file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
(to see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.
Before running the mesher, a number of parameters need to be set in
the \texttt{Mesh\_Par\_file}. This requires a basic understanding
of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
and \citet{KoLiTrSuStSh04}.

The mesher and the solver use UTM coordinates internally, therefore
you need to define the zone number for the UTM projection (e.g., zone
11 for Los Angeles). Use decimal values for latitude and longitude
(no minutes/seconds). These values are approximate; the mesher will
round them off to define a square mesh in UTM coordinates. When running
benchmarks on rectangular models, turn the UTM projection off by using
the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION},
in which case all `longitude' parameters simply refer to the $x$~axis, and
all `latitude' parameters simply refer to the $y$~axis. To run the
mesher for a global simulation, the following parameters need to be
set in the \texttt{\small Mesh\_Par\_file}:
%
\begin{description}
\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
for South).
\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
for West).
\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
which your model resides, only valid when
\texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
Use a negative zone number for the Southern hemisphere:
the Northern hemisphere corresponds to zones +1 to +60,
the Southern hemisphere to zones -1 to -60.

@@

Finally, depending on your system, you might need to provide a file
that tells MPI what compute nodes to use for the simulations. The
file must have a number of entries (one entry per line) at least equal
to the number of processors needed for the run. A sample file is provided
in the file \texttt{mymachines}. This file is not used by the mesher
or solver, but is required by the \texttt{go\_mesher} and \texttt{go\_solver}
default job submission scripts. See Chapter~\ref{cha:Scheduler}
for information about running the code on a system with a scheduler,
e.g., LSF.

Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
and have compiled the mesher, you are ready to launch it! This is
most easily accomplished based upon the \texttt{go\_mesher} script.
When you run on a PC cluster, the script assumes that the nodes are
named n001, n002, etc. If this is not the case, change the \texttt{tr
-d `n'} line in the script. You may also need to edit the last command
at the end of the script that invokes the \texttt{mpirun} command.
See Chapter~\ref{cha:Scheduler} for information about running the
code on a system with a scheduler, e.g., LSF.

Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
in \texttt{output\_mesher.txt}; this file provides lots of details
about the mesh that was generated. Please note that the mesher suggests
a time step \texttt{DT} to run the solver with. The mesher output
file also contains a table about the quality of the mesh to indicate
possible problems with the distortions of elements. Alternatively,
output can be directed to the screen instead by uncommenting a line
in \texttt{constants.h}:
\begin{verbatim}
! uncomment this to write messages to the screen
! integer, parameter :: IMAIN = ISTANDARD_OUTPUT
\end{verbatim}
To control the quality of the mesh, check the standard output (either
on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
and analyze the skewness histogram of mesh elements to make sure no
element has a skewness above approximately 0.75, otherwise the mesh
is of poor quality (and if even a single element has a skewness value
above 0.80, then you must definitely improve the mesh). To draw the
skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.

>>>>

\noindent
When you are ready to compile, in the main directory type
{\small
\begin{verbatim}
  make xdecompose_mesh
\end{verbatim}
}
\noindent
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
should be produced.

@@

In case you successfully ran the configuration script, you are also
ready to compile the internal mesher. This is an alternative to CUBIT
for the mesh generation of relatively simple models.\\

\begin{figure}[htbp]
\begin{centering}
\includegraphics[scale=0.5]{figures/socal_map_mpi.pdf}
\par
\end{centering}
\caption{For parallel computing purposes,
the model block is subdivided in $\nprocxi\times\nproceta$ slices
of elements. In this example we use $5^{2}=25$ processors. }
\label{fig:For-parallel-computing}
\end{figure}

\noindent
In the main directory, type
{\small
\begin{verbatim}
  make xmeshfem3D
\end{verbatim}
}
\noindent
If all paths and flags have been set correctly, the mesher should
now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
note that \texttt{xmeshfem3D} must be called directly from the main
directory, as most of the binaries of the package.\\

Input for the mesh generation program is provided through the parameter
file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
(To see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.)
Before running the mesher, a number of parameters need to be set in
the \texttt{Mesh\_Par\_file}. This requires a basic understanding
of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
and \citet{KoLiTrSuStSh04}.\\

The mesher and the solver use UTM coordinates internally, therefore
you need to define the zone number for the UTM projection (e.g., zone
11 for Los Angeles). Use decimal values for latitude and longitude
(no minutes/seconds). These values are approximate; the mesher will
round them off to define a square mesh in UTM coordinates. When running
benchmarks on rectangular models, turn the UTM projection off by using
the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION},
in which case all `longitude' parameters simply refer to the $x$~axis, and
all `latitude' parameters simply refer to the $y$~axis.\\

\noindent
To run the
internal mesher for a regional simulation, the following parameters need to be
set in the \texttt{\small Mesh\_Par\_file}:
%
\begin{description}
\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
for South).
\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
for West).
\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
which your model resides, only valid when
\texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
Use a negative zone number for the Southern hemisphere:
the Northern hemisphere corresponds to zones +1 to +60,
the Southern hemisphere to zones -1 to -60.

@@

Now that you have set the appropriate parameters in the \texttt{Mesh\_Par\_file}
and have compiled the mesher, you are ready to run it!
\footnote{
Depending on your system, please check how to run executables with MPI.
To run the mesher on your cluster, we provide some example scripts for you to modify accordingly in \texttt{utils/Cluster/}.
For some installations, you might need to provide a file
that tells MPI what compute nodes to use for the simulations. In this case, the
file must have a number of entries (one entry per line) at least equal
to the number of processors needed for the run.
A sample file name is \texttt{mymachines}. This file is not used by the mesher
or solver, but is required by the \texttt{go\_mesher***} and \texttt{go\_solver**}
default job submission scripts provided in directory \texttt{utils/Cluster/}.
To run the mesher, it is most easily accomplished based upon the \texttt{go\_mesher***} script.
When you run on a PC cluster, the script assumes that the nodes are
named n001, n002, etc. If this is not the case, change the \texttt{tr
-d `n'} line in the script. You may also need to edit the last command
at the end of the script that invokes the \texttt{mpirun} command.
See Chapter~\ref{cha:Scheduler}
for information about running the code on a system with a scheduler,
e.g., LSF.
}
Mesher output is provided in the \texttt{OUTPUT\_FILES/} directory
in \texttt{output\_mesher.txt}; this file provides lots of details
about the mesh that was generated.\\

Please note that the mesher suggests
a time step \texttt{DT} to run the solver with. The mesher output
file also contains a table about the quality of the mesh to indicate
possible problems with the distortions of elements. Alternatively,
output can be directed to the screen instead by uncommenting a line
in \texttt{constants.h}:
{\small
\begin{verbatim}
! uncomment this to write messages to the screen
! integer, parameter :: IMAIN = ISTANDARD_OUTPUT
\end{verbatim}}

To control the quality of the mesh, check the standard output (either
on the screen or in the \texttt{OUTPUT\_FILES/} directory in \texttt{output\_mesher.txt})
and analyze the skewness histogram of mesh elements to make sure no
element has a skewness above approximately 0.75, otherwise the mesh
is of poor quality (and if even a single element has a skewness value
above 0.80, then you must definitely improve the mesh). To draw the
skewness histogram on the screen, you could type
{\small \begin{verbatim}
  gnuplot plot_mesh_quality_histogram.gnu
\end{verbatim}}
